---
title: "Locations_updates"
format: html
---

This is the new file for finding all of the plant locations 

#loads in plant locations and gets location coordinates 
```{r}
p_location_data <- read.csv("~/Downloads/Locations_updated.csv")

library(dplyr)
library(ggmap)
library(tidyr)
library(purrr)

register_google(key = "AIzaSyCv5KctcPq9IS4v-etomxY__vI0l8ZBd-M" ) 

p_location_data2 <- p_location_data %>%
  mutate(geocode_result = purrr::map(Address, ~geocode(.x, source = "google"))) %>%
  unnest(geocode_result)

```

#deleted old lat and lon column 
```{r}
p_location_data2 <- p_location_data2 %>%
  select(-c(Lon,Lat))
```

#created for visualizing datapoints so it does not change the type of data p_location_data2 is 
```{r}
p_location_vis <- p_location_data2
```

# creates a data vis of all of the points 
```{r}
# load packages
library(data.table)
library(sp)
library(sf)
library(ggplot2)
library(plotly) 
library(RColorBrewer)

coordinates(p_location_vis) = c("lon", "lat")
crs.geo1 = CRS("+proj=longlat")
proj4string(p_location_vis) = crs.geo1
plot(p_location_vis, pch = 20, col = "steelblue")

```






# points map 
```{r}
library(plotly)


colors <- as.factor(p_location_data2$Company)

# Plot US map with points
plot_geo(p_location_data2, locationmode = 'USA-states') %>%
  add_markers(
    x = ~lon,
    y = ~lat,
    color = ~Company,
    marker = list(
      size = 7,   # adjusts size of the dots 
      line = list(width = 1, color = 'black')  # adjusts the outline of the points 
    )
  ) %>%
  layout(
    title = "PFAS Production Hotspots Across America",
    geo = list(
      scope = 'usa',           
      projection = list(type = 'albers usa'), 
      showland = TRUE,
     landcolor = 'rgb(240,240,240)',
    subunitcolor = 'rgb(200,200,200)',
     countrycolor = 'rgb(200,200,200)'
    ),
    legend = list(x = 1.05, y = 1)  #adjusts legend position
  )

```


# better filled map 
```{r}
library(ggplot2)
library(sf)
library(tigris)
library(dplyr)
library(plotly)
library(grid)
library(gridExtra)
library(jpeg)
library(png)

options(tigris_use_cache = TRUE)

# Get county boundaries for the entire US
counties <- counties(cb = TRUE, resolution = "20m", year = 2021)

# Get state boundaries
states <- states(cb = TRUE, resolution = "20m", year = 2021)

# Filter to continental US only
continental_counties <- counties %>%
  filter(!STATEFP %in% c("02", "15", "60", "66", "69", "72", "78"))

continental_states <- states %>%
  filter(!STATEFP %in% c("02", "15", "60", "66", "69", "72", "78"))

# Convert your plant location data to sf object
plants_sf <- st_as_sf(p_location_data2, 
                      coords = c("lon", "lat"), 
                      crs = 4326)

# Ensure CRS match
continental_counties <- st_transform(continental_counties, crs = 4326)
continental_states <- st_transform(continental_states, crs = 4326)

# Spatial join
plants_with_counties <- st_join(plants_sf, continental_counties)

# Get unique counties with plants
counties_with_plants <- plants_with_counties %>%
  st_drop_geometry() %>%
  select(GEOID, Company) %>%
  distinct()

# Join back to county polygons
counties_filled <- continental_counties %>%
  left_join(counties_with_plants, by = "GEOID")

# Create ggplot
p <- ggplot() +
  geom_sf(data = counties_filled, 
          aes(fill = Company), 
          color = "black", 
          size = 0.05) +
  geom_sf(data = continental_states, 
          fill = NA, 
          color = "black", 
          size = 0.25) +
  scale_fill_discrete(na.value = "#f5f1e7") +
  coord_sf(xlim = c(-125, -66), ylim = c(24, 50), expand = FALSE) +
  theme_void() +
  labs(title = "PFAS Production Hotspots Across America",
       fill = "Company") +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 16, hjust = 0.5, margin = margin(b = 10))
  )

# Load the JPG image
img <- readPNG("N.png")  # Change to your image path
img_grob <- rasterGrob(img, interpolate = TRUE)

# Combine map and image
combined <- gridExtra::grid.arrange(p, img_grob, nrow = 2, heights = c(6, 1))

# Display
plot(combined)



```



# plotly map 
```{r}
library(ggplot2)
library(sf)
library(tigris)
library(dplyr)
library(plotly)
library(grid)
library(gridExtra)
library(jpeg)
library(htmltools)

# Set options to avoid unnecessary messages
options(tigris_use_cache = TRUE)

# Get county boundaries for the entire US
counties <- counties(cb = TRUE, resolution = "20m", year = 2021)

# Get state boundaries
states <- states(cb = TRUE, resolution = "20m", year = 2021)

# Filter to continental US only
continental_counties <- counties %>%
  filter(!STATEFP %in% c("02", "15", "60", "66", "69", "72", "78"))

continental_states <- states %>%
  filter(!STATEFP %in% c("02", "15", "60", "66", "69", "72", "78"))

# Convert your plant location data to sf object
plants_sf <- st_as_sf(p_location_data2, 
                      coords = c("lon", "lat"), 
                      crs = 4326)

# Ensure CRS match
continental_counties <- st_transform(continental_counties, crs = 4326)
continental_states <- st_transform(continental_states, crs = 4326)

# Spatial join
plants_with_counties <- st_join(plants_sf, continental_counties)

# Get unique counties with plants
counties_with_plants <- plants_with_counties %>%
  st_drop_geometry() %>%
  select(GEOID, Company) %>%
  distinct()

# Join back to county polygons
counties_filled <- continental_counties %>%
  left_join(counties_with_plants, by = "GEOID")

# Create ggplot
p <- ggplot() +
  geom_sf(data = counties_filled, 
          aes(fill = Company), 
          color = "black", 
          size = 0.05) +
  geom_sf(data = continental_states, 
          fill = NA, 
          color = "black", 
          size = 0.25) +
  scale_fill_discrete(na.value = "#f5f1e7") +
  coord_sf(xlim = c(-125, -66 ), ylim = c(24, 50), expand = FALSE) + # -66
  theme_void() +
  labs(title = "PFAS Production Hotspots Across America",
       fill = "Company") +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 16, hjust = 0.5, margin = margin(b = 10))
  )

# Convert to plotly
plotly_map <- ggplotly(p, tooltip = "fill")

# Load the JPG image and convert to base64
img_path <- "N.png"  # Change to your image path
img_base64 <- base64enc::base64encode(img_path)

# Add image as an annotation at the bottom
plotly_with_img <- plotly_map %>%
  layout(
    images = list(
      list(
        source = paste0("data:image/png;base64,", img_base64),
        xref = "paper",
        yref = "paper",
        x = .56,
        y = -0.15,
        sizex = 1,
        sizey = 0.2,
        xanchor = "center",
        yanchor = "top"
      )
    ),
    margin = list(b = 125)  # Add bottom margin for image
  )

# Display
plotly_with_img


```










































#getting military installation data 

```{r}
library(dplyr)
fire_sites <- read_csv("~/Downloads/List-of-military-fire-and-crash-training-sites-2014 (1).csv")
pfas_progress <- read_csv("~/Downloads/PROGRESS-AT-INSTALLATIONS-BEING-ASSESSED-FOR-PFAS-USE-OR-POTENTIAL-RELEASE.csv")
pfas_prog_internal_clean <- read.csv("pfas_progress_june_2021.csv")
#fire_sites_clean <- fire_sites %>%
  
#pfas_progress_clean <- pfas_progress %>%
```







































Building Plotly map through ggplot with counties 
```{r}
library(tidycensus)
# installing api key 
census_api_key("29a0b950ad6549f8dc16707914d212045d3b5f3a", install = TRUE)
```

```{r}
library(plotly)
library(tidycensus)
library(tidyverse)
library(sf)
library(geojsonsf)

# Your existing code for census data
#census_api_key("29a0b950ad6549f8dc16707914d212045d3b5f3a", install = TRUE)

# Get County Data with population and calculate density
county_data <- get_acs(
  geography = "county",
  variables = "B01003_001",  # Total population
  year = 2022,
  survey = "acs5",
  geometry = TRUE,
  keep_geo_vars = TRUE 
) %>%
  rename(population = estimate) %>%
  # Calculate population density (people per square mile)
  mutate(
    area_sq_miles = as.numeric(st_area(geometry)) / 2589988.11, # Convert sq meters to sq miles
    density = population / area_sq_miles
  )

# Convert to GeoJSON for Plotly
county_geojson <- geojson_sf(county_data)

# Your original point data preparation
# Convert company to factor
colors <- as.factor(p_location_data2$Company)

# Create the combined map
plot_geo() %>%
  # Add county density choropleth
  add_trace(
    data = county_data,
    type = "choropleth",
    geojson = county_geojson,
    locations = ~GEOID,
    z = ~density,
    colorscale = "Blues",
    featureidkey = "properties.GEOID",
    showscale = TRUE,
    name = "Population Density",
    hovertemplate = paste(
      "<b>%{text}</b><br>",
      "Density: %{z:.1f} people/sq mi<extra></extra>"
    ),
    text = ~paste(NAME, "County")
  ) %>%
  # Add company points
  add_markers(
    data = p_location_data2,
    x = ~lon,
    y = ~lat,
    color = ~Company,
    marker = list(
      size = 7,
      line = list(width = 1, color = 'black')
    ),
    name = "Company Locations",
    hovertemplate = paste(
      "<b>%{text}</b><br>",
      "Company: %{marker.color}<extra></extra>"
    ),
    text = ~paste("Location:", ifelse("Location" %in% names(p_location_data2), Location, "Unknown"))
  ) %>%
  layout(
    geo = list(
      scope = 'usa',
      projection = list(type = 'albers usa'),
      showland = TRUE,
      landcolor = 'rgb(240,240,240)',
      subunitcolor = 'rgb(200,200,200)',
      countrycolor = 'rgb(200,200,200)'
    ),
    legend = list(
      x = 1.05,
      y = 1,
      title = list(text = "Companies")
    ),
    title = "Company Locations Overlaid on Population Density"
  )
```




