---
title: "3M_location_data"
format: html
---

################################
YOU CAN NOT RUN THIS ALL OF THE FILES GOT MESSED UP AND REVERTED!!!
IT WILL RUIN EVERYTHING!!!!
################################



# reading in 3M locations 
```{r}
location_data_3M_update <- read.csv("~/Downloads/Microplastics_site R.csv",stringsAsFactors = FALSE)
library(dplyr)
library(stringr)
```
# zip code manipulation 3M locations 
```{r}
location_data_3M_update <- location_data_3M_update %>% 
  mutate(zip.code = gsub("[^0-9]", "", zip.code)) %>% # takes out non numbers 
  mutate(zip.code = substr(zip.code, 1, 5)) # only includes first 5 numbers 


```

add leading 0 to zip codes to 3M addresses 
```{r}
location_data_3M_update$zip.code[9] <- "01824"
location_data_3M_update$zip.code[19] <- "08822"
```

#changing data set 
```{r}
library(dplyr)

# Join the datasets and add full_address as "Address" column
location_data_3M_update <- location_data_3M_update %>%
  left_join(
    location_data_3M %>% select(zip.code, full_address), 
    by = "zip.code"
  ) %>%
  rename(Address = full_address)

# Alternative approach: If you want to be more explicit about column selection
location_data_3M_update <- location_data_3M_update %>%
  left_join(
    location_data_3M %>% 
      select(zip.code, full_address) %>%
      rename(Address = full_address), 
    by = "zip.code"
  )

# If you want to handle cases where zip codes might not match perfectly
# and want to see which ones didn't match:
location_data_3M_update <- location_data_3M_update %>%
  left_join(
    location_data_3M %>% 
      select(zip.code, full_address) %>%
      distinct(zip.code, .keep_all = TRUE),  # Remove any duplicate zip codes
    by = "zip.code"
  ) %>%
  rename(Address = full_address)

# Check for any missing addresses after the join
missing_addresses <- sum(is.na(location_data_3M_update$Address))
cat("Number of records without matching addresses:", missing_addresses, "\n")

# Display summary of the join operation
cat("Total records in updated dataset:", nrow(location_data_3M_update), "\n")
cat("Records with addresses:", sum(!is.na(location_data_3M_update$Address)), "\n")

# Show sample of the updated data
head(location_data_3M_update %>% select(zip.code, Address))



```


#getting lat and lon of 3M locations 
```{r}
library(dplyr)
library(ggmap)
library(tidyr)
library(purrr)


# Update row 37 of the Address column
location_data_3M_update[37, "Address"] <- "19901 Nordhoff Street"

# Register Google API
register_google(key = "AIzaSyCv5KctcPq9IS4v-etomxY__vI0l8ZBd-M")

# Solution: Filter out NA values before geocoding
location_data_3M_update <- location_data_3M_update %>%
  # Filter out rows where Address is NA or empty
  filter(!is.na(Address) & trimws(Address) != "") %>%
  # Now geocode the valid addresses
  mutate(geocode_result = purrr::map(Address, ~geocode(.x, source = "google"))) %>%
  unnest(geocode_result)

# Alternative: Use a safer geocoding function that handles NA values
safe_geocode <- function(address) {
  # Check if address is NA or empty
  if(is.na(address) || trimws(address) == "") {
    return(data.frame(lon = NA, lat = NA))
  }
  
  tryCatch({
    Sys.sleep(0.1)  # Rate limiting
    result <- geocode(address, source = "google")
    return(result)
  }, error = function(e) {
    message(paste("Geocoding failed for:", substr(address, 1, 50)))
    return(data.frame(lon = NA, lat = NA))
  })
}

# Use the safe geocoding function (keeps all rows, geocodes valid ones)
location_data_3M_update <- location_data_3M_update %>%
  mutate(geocode_result = purrr::map(Address, safe_geocode)) %>%
  unnest(geocode_result)

# If you want to see which rows had NA addresses before filtering
na_addresses <- location_data_3M_update %>%
  mutate(row_number = row_number()) %>%
  filter(is.na(Address) | trimws(Address) == "") %>%
  select(row_number, Address)

if(nrow(na_addresses) > 0) {
  cat("Rows with missing addresses:\n")
  print(na_addresses)
}

```


Loading in solvay data
```{r}
location_data_solvay <- read.csv("~/Downloads/Microplastics site R Solvay .csv")
```


```{r}
library(dplyr)
library(ggmap)
library(tidyr)
library(purrr)

register_google(key = "AIzaSyCv5KctcPq9IS4v-etomxY__vI0l8ZBd-M" ) 

location_data_solvay <- location_data_solvay %>%
  mutate(geocode_result = purrr::map(Address, ~geocode(.x, source = "google"))) %>%
  unnest(geocode_result)
```


load in daikin data 
```{r}
location_data_daikin <- read.csv("~/Downloads/Microplastics site R Daikin.csv")
```

```{r}
library(dplyr)
library(ggmap)
library(tidyr)
library(purrr)

register_google(key = "AIzaSyCv5KctcPq9IS4v-etomxY__vI0l8ZBd-M" ) 

location_data_daikin <- location_data_daikin %>%
  mutate(geocode_result = purrr::map(Address, ~geocode(.x, source = "google"))) %>%
  unnest(geocode_result)
```

# AGC data 
```{r}
location_data_AGC <- read.csv("~/Downloads/Microplastics site R AGC.csv")
```

```{r}
library(dplyr)
library(ggmap)
library(tidyr)
library(purrr)

register_google(key = "AIzaSyCv5KctcPq9IS4v-etomxY__vI0l8ZBd-M" ) 

location_data_AGC <- location_data_AGC %>%
  mutate(geocode_result = purrr::map(Address, ~geocode(.x, source = "google"))) %>%
  unnest(geocode_result)
```

#Arkema Data 
```{r}
location_data_Arkema <- read.csv("~/Downloads/Microplastics site R Arkema.csv")
```

```{r}
library(dplyr)
library(ggmap)
library(tidyr)
library(purrr)

register_google(key = "AIzaSyCv5KctcPq9IS4v-etomxY__vI0l8ZBd-M" ) 

location_data_Arkema <- location_data_Arkema %>%
  mutate(geocode_result = purrr::map(Address, ~geocode(.x, source = "google"))) %>%
  unnest(geocode_result)

```



# merging together datasets 
```{r}
library(dplyr)

# Combine all dataframes with company identifier
Full_location_dataset  <- bind_rows(
  location_data_3M_update %>% select(city, state, lon, lat) %>% mutate(company = "3M"),
  location_data_AGC %>% select(city, state, lon, lat) %>% mutate(company = "AGC"),
  location_data_Arkema %>% select(city, state, lon, lat) %>% mutate(company = "Arkema"),
  location_data_daikin %>% select(city, state, lon, lat) %>% mutate(company = "Daikin"),
  location_data_solvay %>% select(city, state, lon, lat) %>% mutate(company = "Solvay")
) %>%
  # Remove rows with missing coordinates
  filter(!is.na(lon), !is.na(lat)) %>%
  # Reorder columns to put company first
  select(company, city, state, lon, lat)

```


Not needed just wanted to look at it 
#visualizing plant locations 
```{r}
# load packages
library(data.table)
library(sp)
library(sf)
library(ggplot2)
library(plotly) 


# make spacial coordinates (from youtube video)
coordinates(Full_location_dataset) = c("lon", "lat")
crs.geo1 = CRS("+proj=longlat")
proj4string(Full_location_dataset) = crs.geo1
plot(Full_location_dataset, pch = 20, col = "steelblue")
```






